---
title: "Excelerate"
author: "Zelibe_Samuel"
date: "2024-09-26"
output:
  pdf_document: default
  html_document: default
---

## Installing and loading packages
```{r}
install.packages("tidyverse")
install.packages("lubridate")
install.packages("ggplot2")
install.packages("janitor")
install.packages("ggalt")
```

```{r}
library(tidyverse)
library(lubridate)
library(ggplot2)
library(janitor)
library(readr)
library(tidyr)
library(readr)
library(dplyr)
library(ggalt)
```

## Import and Rename Data files

```{r}
Owdf <- read_csv("/cloud/project/Opportunity Data Cleaned.csv")
Udf <- read_csv("/cloud/project/UserData (2) Cleaned.csv")
```

## Checking The Number Of Rows, Data Dimension and The First Six Rows 

```{r}
nrow(Owdf)
nrow (Udf)
```
```{r}
dim(Owdf)
dim(Udf)
```
```{r}
head(Owdf)
head(Udf)
```

```{r}
Owdf %>% distinct(`Profile Id`)
unique_profiles_with_opportunities <- Owdf %>% distinct(`Profile Id`) %>% 
nrow()
```
```{r}
total_signed_up <- 20321

chart_data <- data.frame(
  Total_Signed_Up = c(total_signed_up),
  Unique_Profiles_with_Opportunities = c(unique_profiles_with_opportunities))

ggplot(chart_data, aes(x = Total_Signed_Up, y = Unique_Profiles_with_Opportunities)) +
  geom_point(size = 4, color = "blue") +
  geom_text(aes(label = paste("Total Signed Up:", Total_Signed_Up, "\nUnique Profiles with Opportunities:", Unique_Profiles_with_Opportunities)), vjust = -1) +
  labs(title = "Platform Engagement", x = "Total Signed Up", y = "Unique Profiles with Opportunities")
```
##  Top 10 countries learners have signed up from
```{r}
length(unique(Owdf$Country))
sort(table(Owdf$Country), decreasing = TRUE)[1:10]
```

```{r}
country_counts <- table(Owdf$Country)
top_10_countries <- names(sort(country_counts, decreasing = TRUE)[1:10])
filtered_data <- Owdf[Owdf$Country %in% top_10_countries, ]

ggplot(filtered_data, aes(x = Country)) +
  geom_bar() +
  labs(title = "Top 10 Countries", x = "Country", y = "Frequency") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Cities in the US learners have signed up from
```{r}
Owdf_USA <- Owd %>% filter(Country == "United States") %>% arrange(desc(City))
Owdf_USA %>% distinct(City)
```
```{r}
unique_cities <- Owdf_USA %>% distinct(City)

ggplot(unique_cities, aes(x = City)) +
  geom_bar() +
  labs(title = "US Cities in Owdf", x = "City", y = "Frequency") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Most popular opportunity learners have signed up for

```{r}
most_popular_opportunity <- names(which.max(table(Owdf$`Opportunity Name`)))

chart_data <- data.frame(
  Opportunity = c(most_popular_opportunity, "Others"),
  Count = c(table(Owdf$`Opportunity Name`)[most_popular_opportunity],
            sum(table(Owdf$`Opportunity Name`)) - table(Owdf$`Opportunity Name`)[most_popular_opportunity]))

ggplot(chart_data, aes(x = "", y = Count, fill = Opportunity)) +
  geom_bar(stat = "identity", color = "white") +
  coord_polar(theta = "y", start = 0) +
  labs(title = "Most Popular Opportunity") +
  theme_void()
```

## Most popular opportunity learners have completed

```{r}

Owdf_Nam <- Owdf %>% filter(`Status Description` == "Rewards Award") %>% arrange(desc(`Opportunity Name`))
most_popular_opportunity <- names(which.max(table(Owdf_Nam$`Opportunity Name`)))
```

```{r}
chart_data <- data.frame(
  Opportunity = c(most_popular_opportunity, "All Other Opportunities"),
  Count = c(table(Owdf_Nam$`Opportunity Name`)[most_popular_opportunity],
            sum(table(Owdf_Nam$`Opportunity Name`)) - table(Owdf_Nam$`Opportunity Name`)[most_popular_opportunity])
)

ggplot(chart_data, aes(x = Opportunity, y = Count)) +
  geom_lollipop(color = "skyblue") +
  labs(title = "Most Popular Completed Opportunity", x = "Opportunity", y = "Count")
```

## The demographic breakdown gender of those who have signed up and completed

```{r}

Owdf_Nam <- Owdf %>% filter(`Status Description` == "Rewards Award") %>% arrange(desc(`Opportunity Name`))
gender_counts <- Owdf_Nam %>% group_by(Gender) %>% summarise(count = n_distinct(`Profile Id`))
```

```{r}
ggplot(gender_counts, aes(x = Gender, y = count)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Gender Demographic Breakdown", x = "Gender", y = "Count")
```

## the demographic breakdown student status, etc of those who have signed up and completed

```{r}
Owdf_Nams<- Owdf %>% filter(`Status Description` == "Rewards Award") %>% arrange(desc(`Opportunity Name`))
student_status_count <- Owdf_Nam %>% group_by(`Current Student Status`) %>%
  summarise(count = n_distinct(`Profile Id`))
```

```{r}
ggplot(student_status_count, aes(x = `Current Student Status`, y = count)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Distribution of Student Status Counts", x = "Current Student Status", y = "Count")
```

## The most gained skills on Excelerate

```{r}
Owdf_Nam <- Owdf %>% filter(`Status Description` == "Rewards Award") %>% arrange(desc(`Opportunity Name`))
skill_counts <- Owdf_Nam %>% group_by(`Skills Earned`) %>% summarise(count = n_distinct(`Profile Id`))

skill_counts_sorted <- skill_counts[order(-skill_counts$count), ]
```


## The total scholarship awarded and through which opportunities

```{r}
Owdf_Nam <- Owdf %>% filter(`Status Description` == "Rewards Award") %>% arrange(desc(`Opportunity Name`))

opportunity_rewards <- Owdf_Nam %>%
  group_by(`Opportunity Name`) %>%
  summarize(`sum_reward` = sum(`Reward Amount`))
print(opportunity_rewards)
```

```{r}
most_scholarship_opportunity <- opportunity_rewards %>%
  arrange(desc(`sum_reward`)) %>%
  head(1)
print(most_scholarship_opportunity)
```

```{r}
ggplot(opportunity_rewards, aes(x = reorder(`Opportunity Name`, `sum_reward`), y = `sum_reward`)) +
  geom_lollipop(color = "orange") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(title = "Diverging Lollipop Chart: Total Rewards by Opportunity",
       x = "Opportunity Name",
       y = "Total Reward Amount") +
  coord_flip() +
  theme_minimal()
```



